---
title: "Kaplan-Meier Survival Analysis with Strategus"
author: "Strategus Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Kaplan-Meier Survival Analysis with Strategus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates how to use the Kaplan-Meier Survival module in Strategus to perform Kaplan-Meier survival analysis on observational data in the OMOP Common Data Model (CDM). The Kaplan-Meier Survival module integrates the [CohortSurvival](https://github.com/darwin-eu/CohortSurvival) package to provide Kaplan-Meier survival analysis capabilities.

## Prerequisites

To use the Kaplan-Meier Survival module, you need to install the CohortSurvival package and its dependencies:

```r
# Install CohortSurvival and its dependencies from GitHub
# Note: These packages require specific versions as specified in the DESCRIPTION file
remotes::install_github("darwin-eu/CDMConnector")
remotes::install_github("darwin-eu/PatientProfiles")
remotes::install_github("darwin-eu/omopgenerics")
remotes::install_github("darwin-eu/visOmopResults")
remotes::install_github("darwin-eu/CohortSurvival")

# Install survival package if not already installed
if (!requireNamespace("survival", quietly = TRUE)) {
  install.packages("survival")
}

# Load required libraries
library(CohortSurvival)
library(CDMConnector)
```

## Setting up the Analysis

First, let's create the analysis specifications for a Kaplan-Meier survival analysis study:

```{r}
library(Strategus)

# Create empty analysis specifications
analysisSpecifications <- createEmptyAnalysisSpecificiations()

# Create Kaplan-Meier survival analysis module specifications
cohortSurvivalModuleSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table",
  strata = list(c("age_group"), c("sex")),
  timeGap = 7,
  followUp = 365,
  minCellCount = 5
)

# Add the Kaplan-Meier survival analysis module to the analysis specifications
analysisSpecifications <- addCohortSurvivalModuleSpecifications(
  analysisSpecifications = analysisSpecifications,
  moduleSpecifications = cohortSurvivalModuleSpecifications
)
```



## Execution Settings

Set up the execution settings for your database:

```{r}
# Create execution settings
executionSettings <- createCdmExecutionSettings(
  workDatabaseSchema = "scratch_schema",
  cdmDatabaseSchema = "cdm_schema",
  cohortTableNames = CohortGenerator::getCohortTableNames(cohortTable = "cohort_table"),
  tempEmulationSchema = NULL,
  workFolder = "path/to/work/folder",
  resultsFolder = "path/to/results/folder",
  logFileName = "cohort_survival_analysis_log.txt",
  minCellCount = 5,
  incremental = FALSE,
  maxCores = 4
)
```

## Running the Analysis

Execute the Kaplan-Meier survival analysis:

```{r}
# Execute the analysis
execute(
  connectionDetails = connectionDetails,
  analysisSpecifications = analysisSpecifications,
  executionSettings = executionSettings
)
```

## Results

The Kaplan-Meier Survival module will generate several output files:

- **Kaplan-Meier survival estimates**: CSV files containing survival probabilities and confidence intervals
- **Stratified results**: Results broken down by stratification variables
- **Summary statistics**: Median survival times, restricted mean survival, etc.

## Visualization

After running the analysis, you can use the CohortSurvival package to create visualizations:

```{r}
# Load the results
survivalResults <- CohortGenerator::readCsv("path/to/results/folder/survival_results.csv")

# Create Kaplan-Meier survival plots using the survival results
# Note: You can use ggplot2 or other plotting libraries to visualize the results
library(ggplot2)
ggplot(survivalResults, aes(x = time, y = survival, color = strata_level)) +
  geom_line() +
  facet_wrap(~strata_name) +
  labs(title = "Kaplan-Meier Survival Curves", 
       x = "Time (days)", 
       y = "Survival Probability")
```

## Advanced Configuration

### Custom Strata

You can define custom stratification variables:

```{r}
# Define custom strata
customStrata <- list(
  c("age_group", "sex"),
  c("comorbidity"),
  c("age_group", "comorbidity", "sex")
)

cohortSurvivalModuleSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table",
  strata = customStrata,
  timeGap = 7,
  followUp = 365,
  minCellCount = 5
)
```

### Time Windows

Adjust the time gap and follow-up periods:

```{r}
# Short-term analysis (30 days)
shortTermSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table",
  timeGap = 1,      # Daily intervals
  followUp = 30,    # 30-day follow-up
  minCellCount = 5
)

# Long-term analysis (5 years)
longTermSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table",
  timeGap = 30,     # Monthly intervals
  followUp = 1825,  # 5-year follow-up
  minCellCount = 5
)
```

## Integration with Other Modules

The Kaplan-Meier Survival module can be combined with other Strategus modules:

```{r}
# Add cohort generation
cohortGeneratorSpecifications <- CohortGeneratorModule$new()$createModuleSpecifications()

# Add Kaplan-Meier survival analysis
cohortSurvivalSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table"
)

# Combine in analysis specifications
analysisSpecifications <- createEmptyAnalysisSpecificiations() |>
  addCohortGeneratorModuleSpecifications(cohortGeneratorSpecifications) |>
  addCohortSurvivalModuleSpecifications(cohortSurvivalSpecifications)
```

## Conclusion

The Kaplan-Meier Survival module in Strategus provides a powerful and flexible way to perform Kaplan-Meier survival analysis on observational data. By integrating with the CohortSurvival package, it offers comprehensive Kaplan-Meier survival analysis capabilities with stratification options.

For more information about the CohortSurvival package, visit the [GitHub repository](https://github.com/darwin-eu/CohortSurvival). 