---
title: "Cohort Survival Analysis with Strategus"
author: "Strategus Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cohort Survival Analysis with Strategus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates how to use the Cohort Survival module in Strategus to perform cohort survival analysis on observational data in the OMOP Common Data Model (CDM). The Cohort Survival module integrates the [CohortSurvival](https://github.com/darwin-eu/CohortSurvival) package to provide comprehensive cohort survival analysis capabilities.

## Prerequisites

To use the Cohort Survival module, you need to install the CohortSurvival package:

```r
# Install CohortSurvival from GitHub
remotes::install_github("darwin-eu/CohortSurvival")
```

## Setting up the Analysis

First, let's create the analysis specifications for a cohort survival analysis study:

```{r}
library(Strategus)

# Create empty analysis specifications
analysisSpecifications <- createEmptyAnalysisSpecificiations()

# Create cohort survival analysis module specifications
cohortSurvivalModuleSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  analysisType = "single_event",
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table",
  strata = list(c("age_group"), c("sex")),
  timeGap = 7,
  followUp = 365,
  minCellCount = 5
)

# Add the cohort survival analysis module to the analysis specifications
analysisSpecifications <- addCohortSurvivalModuleSpecifications(
  analysisSpecifications = analysisSpecifications,
  moduleSpecifications = cohortSurvivalModuleSpecifications
)
```

## Competing Risk Analysis

You can also perform competing risk cohort survival analysis:

```{r}
# Create competing risk cohort survival analysis specifications
competingRiskSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  analysisType = "competing_risk",
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table",
  competingOutcomeCohortTable = "cohort_table",
  strata = list(c("age_group")),
  timeGap = 7,
  followUp = 365,
  minCellCount = 5
)

# Add to analysis specifications
analysisSpecifications <- addCohortSurvivalModuleSpecifications(
  analysisSpecifications = analysisSpecifications,
  moduleSpecifications = competingRiskSpecifications
)
```

## Execution Settings

Set up the execution settings for your database:

```{r}
# Create execution settings
executionSettings <- createCdmExecutionSettings(
  workDatabaseSchema = "scratch_schema",
  cdmDatabaseSchema = "cdm_schema",
  cohortTableNames = CohortGenerator::getCohortTableNames(cohortTable = "cohort_table"),
  tempEmulationSchema = NULL,
  workFolder = "path/to/work/folder",
  resultsFolder = "path/to/results/folder",
  logFileName = "cohort_survival_analysis_log.txt",
  minCellCount = 5,
  incremental = FALSE,
  maxCores = 4
)
```

## Running the Analysis

Execute the cohort survival analysis:

```{r}
# Execute the analysis
execute(
  connectionDetails = connectionDetails,
  analysisSpecifications = analysisSpecifications,
  executionSettings = executionSettings
)
```

## Results

The Cohort Survival module will generate several output files:

- **Cohort survival estimates**: CSV files containing survival probabilities and confidence intervals
- **Stratified results**: Results broken down by stratification variables
- **Summary statistics**: Median survival times, restricted mean survival, etc.

## Visualization

After running the analysis, you can use the CohortSurvival package to create visualizations:

```{r}
# Load the results
survivalResults <- CohortSurvival::readSurvivalResults("path/to/results/folder")

# Create survival plots
CohortSurvival::plotSurvival(survivalResults, 
                             colour = "strata_level", 
                             facet = "strata_name")
```

## Advanced Configuration

### Custom Strata

You can define custom stratification variables:

```{r}
# Define custom strata
customStrata <- list(
  c("age_group", "sex"),
  c("comorbidity"),
  c("age_group", "comorbidity", "sex")
)

cohortSurvivalModuleSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  analysisType = "single_event",
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table",
  strata = customStrata,
  timeGap = 7,
  followUp = 365,
  minCellCount = 5
)
```

### Time Windows

Adjust the time gap and follow-up periods:

```{r}
# Short-term analysis (30 days)
shortTermSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  analysisType = "single_event",
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table",
  timeGap = 1,      # Daily intervals
  followUp = 30,    # 30-day follow-up
  minCellCount = 5
)

# Long-term analysis (5 years)
longTermSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  analysisType = "single_event",
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table",
  timeGap = 30,     # Monthly intervals
  followUp = 1825,  # 5-year follow-up
  minCellCount = 5
)
```

## Integration with Other Modules

The Cohort Survival module can be combined with other Strategus modules:

```{r}
# Add cohort generation
cohortGeneratorSpecifications <- CohortGeneratorModule$new()$createModuleSpecifications()

# Add cohort survival analysis
cohortSurvivalSpecifications <- CohortSurvivalModule$new()$createModuleSpecifications(
  analysisType = "single_event",
  targetCohortTable = "cohort_table",
  outcomeCohortTable = "cohort_table"
)

# Combine in analysis specifications
analysisSpecifications <- createEmptyAnalysisSpecificiations() |>
  addCohortGeneratorModuleSpecifications(cohortGeneratorSpecifications) |>
  addCohortSurvivalModuleSpecifications(cohortSurvivalSpecifications)
```

## Conclusion

The Cohort Survival module in Strategus provides a powerful and flexible way to perform cohort survival analysis on observational data. By integrating with the CohortSurvival package, it offers both single event and competing risk cohort survival analysis capabilities with comprehensive stratification options.

For more information about the CohortSurvival package, visit the [GitHub repository](https://github.com/darwin-eu/CohortSurvival). 